% Chapter 1

\chapter{Introduction} % Main chapter title

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1} 

\lhead{Chapter 1. \emph{Introduction}} % This is for the header on each page - perhaps a shortened title
The following assignment is an exam for the master course "MGEW23: Einführung in Bayessche Netze für Geowissenschaftler" in the scope of the master program "Geowissenschaften" at the Universität Potsdam.\\
The purpose of this work is to use Bayesian Networks in an example that resembles simplified questions one would encounter in the assessment of natural hazards. The task is to build ground motion models from synthetic data and quantify their ability to predict values of peak ground acceleration (PGA) given the input variables.\\
In the context of this paper four discrete Bayesian Networks have been learned. These include a causal network, a Naive Bayes network, a constraint-based network using a Grow-Shrink algorithm and a score-based one from a Hill-Climber method. Special care is given to the evaluation of the prediction performance. A mean squared error and a log-likelihood score are computed on a validation set in order to get an estimate of the out-of-sample performance. In a next step the concept of testing on unseen data is expanded to cross-validation and finally an attempt of a bias-variance decomposition is made to further develop this concept and to draw conclusions where potential for improvement lies.\\
The computations were performed using R~\citep{R} in combination with the IDE RStudio~\citep{Rstudio}. For setting up and working with Bayesian Networks the R package bnlearn \citep{bnlearn} was used. \\\\
DISCLAIMER:\\
There are some footnotes throughout this paper. These are personal comments and reflect more of an opinion or thought and are a bit more colloquial than the rest of this document. They are intended as discussion points.